---
title: "Peer_Assignment_1"
author: "Mustafa"
date: "March 31, 2018"
---

# Loading and preprocessing the data
```{r results='hide'}
  library(ggplot2) 
  data <- read.csv("activity.csv")
```

# What is mean total number of steps taken per day?
```{r results='hide'}
  spd <- aggregate(steps ~ date, data, sum, na.rm=TRUE)
```

## 1. Make a histogram of the total number of steps taken each day
```{r results='hide'}
  barplot(spd$steps)   
```

##2. Calculate and report the mean and median total number of steps taken per day
```{r results='hide'}
  smean <- mean(spd$steps)
  smedian <- median(spd$steps)
```
* Mean is: `r smean`
* Median is: `r smedian`


# What is the average daily activity pattern?
```{r results='hide'}
  avg <- tapply(data$steps, data$interval, mean, na.rm=TRUE)   
```

##1. Make a time series plot
```{r results='hide'}
   plot(names(avg), avg, xlab="5-min interval", type="l", ylab="the average number of steps")  
```

##2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r results='hide'}
  maximum_avg <- max(avg)
  maximum_interval <- as.numeric(names(avg)[which(avg==max(avg))])
```
* Maximum steps: `r maximum_avg`
* Maximum is at interval: `r maximum_interval`

# Imputing missing values
##1. Calculate and report the total number of missing values in the dataset
```{r results='hide'}
   nas <- sum(is.na(data$steps))
```
* Total number of NAs is: `r nas`

##2. Devise a strategy for filling in all of the missing values in the dataset

##3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r results='hide'}
  new_data <- data
  new_data$steps[which(is.na(data$steps))] <- as.vector(avg[as.character(data[which(is.na(data$steps)),3])])
```

##4. Make a histogram of the total number of steps taken each day
```{r results='hide'}
  spd2 <- aggregate(steps ~ date, new_data, sum, na.rm=TRUE)
  barplot(spd2$steps)
```

## and calculate and report the mean and median total number of steps taken per day.
```{r results='hide'}
  smean2 <- mean(spd2$steps)
  smedian2 <- median(spd2$steps)
```

* New mean is: `r smean2`
* New median is: `r smedian2`

# Are there differences in activity patterns between weekdays and weekends?

##1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.'
```{r results='hide'}
  new_data$dayType <- ifelse(as.POSIXlt(new_data$date)$wday %in% c(0,6), "weekends","weekdays")
  smedian2 <- median(spd2$steps)
```

##2. Make a panel plot containing a time series plot
```{r results='hide'}
  aggregateData <- aggregate(steps ~ interval + dayType, data=new_data, mean)
  ggplot(aggregateData, aes(interval, steps)) + geom_line() + facet_grid(dayType ~ .) + 
    xlab("5-minute interval") + ylab("avarage number of steps")
```


